---
title: "Day 6 Homework"
output:
  html_document:
    fig_height: 3
    fig_width: 5
    highlight: null
    theme: journal
---

### 1. Think carefully about your data.  Think about something you would like to do multiple times, to subsets of your data.  
Reading in and examining data
```{r}
library(plyr)

bildata <- read.csv("bildataHW6.csv")
CCdata <- read.csv("CCdataHW6.csv")
```

### 2. Write out (in plain English, in the notebook) what you are trying to do
I want to do two things:

1. I want to make sure that the columns in each of my two dataframes (bildata and CCdata) that should be factored are factored.
2. I want to remove the two monolingual and two bilingual participants from both dataframes (CCdata and bildata) who have the lowest accuracy rates.

### 3. Write some "pseudo-code" comments in an R chunk, as an outline
**A. Factoring columns in dataframes CCdata and bildata**
```{r}
# for loop from i = 1 to 9
  # if i < 3
    # then factor column i in bildata dataframe
  # end if statement
  # factor column i in CCdata dataframe
# end for loop
```

**B. Removing the four participants who have the lowest accuracy rates**
```{r}
# use ddply() to create a new dataframe ("accSummary") based on CCdata dataframe that contains mean accuracy rate by subject
# use with() and order() to sort accSummary dataframe by mean accuracy and find IDs corresponding to the two lowest billingual and two lowest monolingual accuracy rates 
# create a vector using as.vector() of these four IDs

# for loop from i = 1 to length of vector of IDs
  # use droplevels(subset()) to remove i ID from CCdata dataframe
  # use droplevels(subset()) to remove i ID from bildata dataframe
# end for loop
```

### 4. Write some (real) code that will do what you want ONCE, on one subset of the data.
**A. Factoring columns in dataframes CCdata and bildata**
```{r}
bildata[,1] <- factor(bildata[,1])
CCdata[,1] <- factor(CCdata[,1])
```

**B. Removing the four participants who have the lowest accuracy rates**
```{r}
accSummary <- ddply(CCdata, c("ID","Lang"), summarise, meanAcc = round(mean(Acc),3), AccSD = round(sd(Acc), 3))

head(accSummary)

accSummary.sorted <- accSummary[with(accSummary, order(meanAcc)), ]

accSummary.sorted
```

The above output indicates that subjects 131, 176, 145, and 121 have the four lowest accuracy rates, so I'll remove 131 from the bildata and CCdata dataframes
```{r}
CCdata.temp <- droplevels(subset(CCdata, ID != "S131"))
bildata.temp <- droplevels(subset(bildata, ID != "S131"))
```

### 5. Copy that code chunk, and convert it into a loop, which will do what you want over all the groups, concatenating the results.
**A. Factoring columns in dataframes CCdata and bildata**  
I want to factor columns 1-2 in bildata and columns 1-9 in CCdata, but I'd like to do both in one for loop, so I'll use an embedded if statement
```{r}
for (i in 1:9 ) {
  if(i < 3) {     
    bildata[,i] <- factor(bildata[,i])
  }
  CCdata[,i] <- factor(CCdata[,i])
}
```

Checking my work
```{r}
for (i in 1:9 ) {
  if(i < 3) {   
    print(is.factor(bildata[,i]))
  }
  print(is.factor(CCdata[,i]))
}  
```

**B. Removing the four participants who have the lowest accuracy rates**
```{r}
accSummary.sorted 

IDs.to.remove <- as.vector(accSummary.sorted[1:4, 1], mode = "character")
IDs.to.remove

for (i in 1:length(IDs.to.remove)) {
  CCdata <- droplevels(subset(CCdata, ID != IDs.to.remove[i])) # remove ID from CCdata
  bildata <- droplevels(subset(bildata, ID != IDs.to.remove[i])) # remove ID from bildata
  }
```

Checking my work
```{r}
unique(CCdata$ID)
length(unique(CCdata$ID)) # should be 60

unique(bildata$ID)
length(unique(bildata$ID)) # should be 30
```

### 6. Copy that loop again, and at a condition (`if`) statement to add some further control about ignoring some groups or some other conditional operation.
**A. Factoring columns in dataframes CCdata and bildata**  
```{r}
for (i in 1:9 ) {
  if(i < 3) {    
    if(is.factor(bildata[,i]) == FALSE) {  # check to see whether current column is already factored; if it isn't, then factor it
      bildata[,i] <- factor(bildata[,i])
    }
  }
  if(is.factor(CCdata[,i]) == FALSE) {   # check to see whether current column is already factored; if it isn't, then factor it
    CCdata[,i] <- factor(CCdata[,i])
  }
}
```

Checking my work
```{r}
for (i in 1:9 ) {
  if(i < 3) {   
    print(is.factor(bildata[,i])) 
  }
  print(is.factor(CCdata[,i]))
}  
```